Seismic parameter estimation and the Canadian crust

### 2013-07-07 paper/draftBP.tex
paper

  Several published regional studies utilizing the methods outlined in section \ref{section:VpVsMethod}, ZK + semblance weighting, provide parameter estimates for $R$ and $H$ for regionally grouped subsets of Canadian seismic stations. Comparisons between estimates for particular seismic stations are made for those values which have a corresponding standard error in $R$ of less than 0.06. This error value best partitions the data between those receiver function stacks with visually recognisable energy along phase moveout curves from those where lack of data or poor quality make these moveout curves difficult to resolve. For each study under comparison, data is reprocessed to use the $V_P$ chosen by the study authors and both correlation and mean difference are calculated (Table \ref{table:comparison}).

@@

\subsection{Regional Bulk Crustal Parameters}
[[ Move from ZK + semblance to Full Grid Search + semblance ]]
  Data from X seismic stations across Canada are analyzed, using only those stations with less than 0.06 standard error. To compensate for the uneven geographic distribution of seismic stations, weights are applied to all crustal parameter estimates before averaging over a given region. Wieghts are calculated by first projecting station locations onto the 2D plane using the Albers equal-area conic projection. A Voronoi diagram is calculated for all projected station locations. The ratio between the Voronoi cell surface encompassing each station and the total area of the convex hull bounding region is computed and used as the weighting value.

@@

  Moving South from the Churchill through the Superior and the Grenville there is an increase in the ZK + semblance computed seismic velocity ratio as well as crustal thickness. The trend in increasing values is also seen in the weighted averaged active source $V_P$ data. Generally the Crust 2.0 values show a slightly thinner crust though the southward thickening trend is noticeble. The trend in the seismic velocity data is not as apparent in the Crust 2.0 data.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{activeSourceComparison}
  \caption{Comparison of $V_P$ estimates from the MB algorithm with active source $V_P$ recordings. Stations selected if active source experiment locations within 1 degree of latitude / longtitude.}
  \label{fig:activeComp}
\end{figure}

@@

  The preliminary interrogation of the data set yields the observation that the bulk Canadian Shield has lower Vp/Vs than anticipated. Previous experiments show crustal averages of 1.77 (Christensen and Mooney, 1995) and 1.78 (Zandt and Ammon, 1995). This compares to computed values of 1.73, 1.74, 1.795 for the Churchill, Superior and Grenville Provinces Vp/Vs ratios respectively.

(MIGHT NOT INCLUDE)  The data bares out earlier results showing Proterozoic crust to have a higher seismic velocity ratio than Archean crust. This follows previous studies on crustal formation that have noted this trend in other regions (Durrheim and Mooney, 1991) . The increased crustal thickness of Proterozoic crust is clearly seen in the active source data while it not visible in data from processed seismic stations.

  Further work on the Bostock-Kumar stacking approach is warranted from a look at the cleanest stations. Before it can be employed in large scale analysis additional denoising methods or alternative deconvolution techniques will need to be investigated to reduce the noise in the data.

>>>>

  Several published regional studies utilizing the methods outlined in section \ref{section:VpVsMethod}, ZK + semblance weighting, provide parameter estimates for $R$ and $H$ for regionally grouped subsets of Canadian seismic stations. Comparisons between estimates for particular seismic stations are made for those values which have a corresponding standard error in $R$ of less than 0.06. This error value best partitions the data between those receiver function stacks with visually recognizable energy along phase moveout curves from those where lack of data or poor quality make these moveout curves difficult to resolve. For each study under comparison, data is reprocessed to use the $V_P$ chosen by the study authors and both correlation and mean difference are calculated (Table \ref{table:comparison}).

@@

\subsection{Regional Bulk Crustal Parameters}
[[ Move from ZK + semblance to Full Grid Search + semblance ]]
  Data from X seismic stations across Canada are analyzed, using only those stations with less than 0.06 standard error. To compensate for the uneven geographic distribution of seismic stations, weights are applied to all crustal parameter estimates before averaging over a given region. Weights are calculated by first projecting station locations onto the 2D plane using the Albers equal-area conic projection. A Voronoi diagram is calculated for all projected station locations. The ratio between the Voronoi cell surface encompassing each station and the total area of the convex hull bounding region is computed and used as the weighting value.

@@

  Moving South from the Churchill through the Superior and the Grenville there is an increase in the ZK + semblance computed seismic velocity ratio as well as crustal thickness. The trend in increasing values is also seen in the weighted averaged active source $V_P$ data. Generally the Crust 2.0 values show a slightly thinner crust though the southward thickening trend is noticeable. The trend in the seismic velocity data is not as apparent in the Crust 2.0 data.

\subsection{Conrad Discontinuity}
Results and figures showing statistics regarding depth of crustal reflectors.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{activeSourceComparison}
  \caption{Comparison of $V_P$ estimates from the MB algorithm with active source $V_P$ recordings. Stations selected if active source experiment locations within 1 degree of latitude / longitude.}
  \label{fig:activeComp}
\end{figure}

@@

  - Comparison Discussion
    - Crustal Thickness
      - Very good correlation in thickness estimates with all three datasets.
    - Vp/Vs
      - Lower correlation with the Vp/Vs values.
        - Expect lower correlation with Vp/Vs since the travel time equations are
          less sensitive to seismic velocities than they are to crustal thickness.
          This means that there will be more uncertainty during the parameter
          search for Vp/Vs and hence higher degree of relative error.
        - ??? Utilizing different sources, applying different deconvolution and
          filtering techniques will therefore have a greater impact on the
          uncertainty of picking Vp/Vs than that of crustal thickness H. ???

   - Explanation of results
    - Canada H and Vp/Vs
      - Lith5.0 (Perry, 2002) estimates for Canadian crustal averages is 38 km. The Crust
        2.0 average is 38km as well, however this is expected since Crust2.0 data for
        Canada is based on the same data as that of Lith5.0
      - ??? Make some comparison to global average ???
